Robustness of MinD oscillation in \emph{Escherichia coli} with diverse cell shapes

### 2014-09-10 paper/paper.tex
more edits to paper Tested everything on quipu.

\section{Naturally Occuring Pill Shaped Cells}
\begin{figure*}
  %\includegraphics[width=\textwidth]{../data/shape-p/3_00-0_50-0_00-0_00-15_00-exact/plots/image-plot}
  \includegraphics[width=\textwidth]{../data/shape-p/3_00-0_50-0_00-0_00-15_00-exact/plots/single-image-plot}
  \includegraphics[width=\textwidth]{../data/shape-p/3_00-0_50-0_00-0_00-15_00-full_array/plots/single-image-plot}
  \caption{\fixme{We should figure out units and use the same levels
      as in the main plot.  I also want to figure out why the two
      plots look different, and make them the same size.  Can we just
      do the gaussian} Contour plot images of the concentration of each protein
    species in a natural pill-shaped bacterium at regular intervals in
    time (one second intervals), with darker regions indicating higher
    concentration. The upper plots shows results from the
    deterministic simulation and the lower shows results from the
    stochastic simulation.  The cells pictured are $4\micron$ in
    length, measured from end to end.  The order of frames is such
    that individual MinD proteins begin at the bottom of the plot (in
    the MinD:ATP state in the cytoplasm), and progress upward until
    they reach the MinE:MinD:ATP membrane-bound complex.  At that
    point, they will spontaneously dissociate into cytoplasmic MinE
    (the top row) and the starting state of cytoplasmic MinD:ADP.  The
    densities plotted are integrated along the axis orthogonal to the
    page, and the color scale is chosen separately for each species
    with black as the maximum value over space and time.  The
    stochastic densities are smeared out over space using a guassians
    approximation developed by Zhang et all~\cite{zhang2007gaussian}}.
  \label{image-p}
\end{figure*}

@@

We test cells of radius $.5\micron$ and of lengths $2\micron$,
$3\micron$, and $5\micron$ measured from the tip of one end cap to the
tip of the other. In each case, we observe in the deterministic
simulations a quick establishement (within a few periods) of extremely
regular oscillatory behavoir that lasts indefinitely.  The stochastic
results show oscillitions that are similarly regular in time, with
periods that nearly match those of the deterministic (they differ
roughly by a factor of .05)\fixme{does this square with the
  correlation data periods?}.

@@

The period increases with cell length, as shown in
Table~\ref{tab:pill-periods}, in agreement with the results of Huang
\emph{et al.}.  The stochastic simulations show a similar pattern.
The MinD maxima occur at the ends of the cell in a back and forth
manner, and with the same approximate period.

@@

\section{Comparison of Mannik's Experimental and Our Stadium Shapes}
\fixme{should I say anything about different/large sizes?}  We have
made movies out of color plots of MinD concentration, integrated along
the flattened axis, in which each frame is seperated in simulation
time by a half second. Our simulations include the two flattened cell
shapes that are reported in Manniks paper, a few other 'irregular'
shapes of our own creation, triangle shaped flattened cells, and a
symetric 'stadium' shaped cell. For each of these shapes we have
varied the size, from roughly twice the dimensions of Mannik's
reported cells down to the smaller sizes refered to in
section~\ref{sec:stability-analysis}.

We have found that the deterministic simulations of the protein
reaction sequence discussed above results in robust oscillatory
behavoir of polar selection and oscillation in not only symmetrical
shapes, such as the wildtype pill shape, but also in very
assymetrical, flatttened cells, such as those studied by Mannik.
After the lower stability limit of
section~\ref{sec:stability-analysis} has been surpassed, this
qualitative behavoir doesn't seem to have size dependence until cells
reach sizes that have more than twice the dimensions of the Mannik
cells.  At this point the robust oscillations, while still evident,
begin to lose their consistanty.  The Mannik reported cells already
seem to be rather large (wild type cells are typically smaller than
those observed my Mannik
\cite{kubitschek1990cell,kubitschek1968linear,mannik2012robustness}),
so we left off simulating cells any larger than twice this size. The
movies show sharp peaks of MinD in specific locatations, typically at
the edges of the cell and in regions of high radius of curvature,
followed by a spreading out of the MinD out into the cell, with no
real bias towards any location, until it aggregates again in a
specific location at the opposite edge of the cell, maximizing there.
The pattern is repeated again and again with surprising consistency,
which can be seen on the right of Figure~\ref{box-mannik}.  This is
the same plotting method as our box plots above, here showing
stochastic and deterministic simulations of cells that are the same
shape and size of one of Mannik's reported cells.

@@

\section{Interpretation of Data}
\section{Conclusion}
No conclusion :(
\bibliography{paper}

>>>>

\begin{figure*}
  %\includegraphics[width=\textwidth]{../data/shape-p/3_00-0_50-0_00-0_00-15_00-exact/plots/image-plot}
  \includegraphics[width=\textwidth]{../data/shape-p/3_00-0_50-0_00-0_00-15_00-exact/plots/single-image-plot}
  \includegraphics[width=\textwidth]{../data/shape-p/3_00-0_50-0_00-0_00-15_00-full_array/plots/single-image-plot}
  \caption{\fixme{We should figure out units and use the same levels
      as in the main plot.  I also want to figure out why the two
      plots look different, and make them the same size.  Can we just
      do the gaussian} Contour plot images of the concentration of each protein
    species in a natural pill-shaped bacterium at regular intervals in
    time (one second intervals), with darker regions indicating higher
    concentration. The upper plots shows results from the
    deterministic simulation and the lower shows results from the
    stochastic simulation.  The cells pictured are $4\micron$ in
    length, measured from end to end.  The order of frames is such
    that individual MinD proteins begin at the bottom of the plot (in
    the MinD:ATP state in the cytoplasm), and progress upward until
    they reach the MinE:MinD:ATP membrane-bound complex.  At that
    point, they will spontaneously dissociate into cytoplasmic MinE
    (the top row) and the starting state of cytoplasmic MinD:ADP.  The
    densities plotted are integrated along the axis orthogonal to the
    page, and the color scale is chosen separately for each species
    with black as the maximum value over space and time.  The
    stochastic densities are smeared out over space using a guassians
    approximation developed by Zhang et all~\cite{zhang2007gaussian}}.
  \label{image-p}
\end{figure*}
\section{Naturally Occuring Pill Shaped Cells}

@@

We test cells of radius $.5\micron$ and of lengths $2\micron$,
$3\micron$, $4\micron$, and $5\micron$ measured from the tip of one
end cap to the tip of the other \fixme{I'm running the 2,3,5 pills
  again to make sure that everything in this paragraph is true}. In
each case, we observe in the deterministic simulations a quick
establishement (within a few periods) of extremely regular oscillatory
behavoir that lasts indefinitely.  The stochastic results show
oscillitions that are similarly regular in time, with periods that
nearly match those of the deterministic (they differ roughly by a
factor of .05) \fixme{does this square with the correlation data
  periods?}. The regularity of the pill shape cell periods is
discussed below, in our analysis of time correlation data \fixme{not
  written yet}.  These periods increase with cell length, as shown in
Table~\ref{tab:pill-periods}, in agreement with the results of Huang
\emph{et al.}.

@@

@@

\fixme{Originally the 'Stability Analysis' section came next.  I would
  like to still mention the results of this, since Huang only did the
  analysis on the cylinders, and we've done it on the flattened shapes
  and shown that it works with those too.  But I don't know where to
  fit in the paragraph or so that it would take.  For now I've just
  put the whole section at the end of the paper.}

\section{Comparison of Mannik's Experimental and Our Stadium Shapes}

We have made and analyzed movies out of color plots of MinD
concentration that are constructed as those in Figure~\ref{image-p},
integrated along the flattened axis, in which each frame is seperated
in simulation time by a half second. Figure~\ref{randst-plot-ave}
illustrates what is seen in these movies. In his work Mannik presents
arrow plots in which the arrow heads show the location of sequential
MinD maxima within the cell. His plots for his two abnormal cell
shapes are shown in the upper 'experiment' section of
Figure~\ref{randst-plot-ave}.  Following this, we have created plots
that show maxima that are global in space and local in time.  Of these
maxima we only show those which have greater concentration than a
chosen cutoff concentration, and also ignore maxima that appear
temporalily and spatially adjacent to preceding maxima, in order to
stop single spikes from appearing as many adjecent ones.  The overal
effect is to accurately describe what is seen in our movies.

We have found that the deterministic simulations of the protein
reaction sequence discussed above results in robust oscillatory
behavoir of polar selection and oscillation in not only symmetrical
shapes, such as the wildtype pill shape, but also in very
assymetrical, flatttened cells, such as those studied by Mannik.
After the lower stability limit of
section~\ref{sec:stability-analysis} \fixme{this needs to be covered}
has been surpassed, this qualitative behavoir doesn't seem to have
size dependence until cells reach sizes that have more than twice the
dimensions of the Mannik cells.  At this point the robust
oscillations, while still evident, begin to lose their consistanty.
The Mannik reported cells already seem to be rather large (wild type
cells are typically smaller than those observed my Mannik
\cite{kubitschek1990cell,kubitschek1968linear,mannik2012robustness}),
so we left off simulating cells any larger than twice this size.

The movies show sharp peaks of MinD in specific locatations, typically at
the edges of the cell and in regions of high radius of curvature,
followed by a spreading out of the MinD out into the cell, with no
real bias towards any location, until it aggregates again in a
specific location at the opposite edge of the cell, maximizing there.
The pattern is repeated again and again with surprising consistency,
which can be seen on the right of Figure~\ref{box-mannik}.  This is
the same plotting method as our box plots above, here showing
stochastic and deterministic simulations of cells that are the same
shape and size of one of Mannik's reported cells.

@@

\fixme{should I say anything about different/large sizes?}

\section{Conclusion}
No conclusion :(
\bibliography{paper}
